The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-07 paper/paper.tex
paper: corrections

\newcommandx{\deps}[2][1,2]{\depsp{}[][][][#1][#2][][]}
\newcommandx{\fulldeps}[1][1]{\depsp{full}[][][][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\depsp{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohs}[1][1]{\depsp{fullcohs}[][][][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\depsp{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscoht}[1][1]{\depsp{fullcoht}[][][][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\depstypep{}[][][][#1][#2][][]}
\newcommandx{\fulldepstype}[2][1,2]{\depstypep{full}[][][][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\depstypep{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohstype}[1][1]{\depstypep{fullcohs}[][][][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\depstypep{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscohttype}[1][1]{\depstypep{fullcoht}[][][][#1][][][]}

@@

Assuming $\framep[n][\lbrack 0,p \rbrack]$ for $p \leq n$ only requires to define the type of frames, which are just finite products of $\U$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][\lbrack 0,p \rbrack] &  & :      & \U                                               \\
    \frametype[n][]                    &  & \defeq & \unittype                                        \\
    \frametype[n][\lbrack 0,p \rbrack] &  & \defeq & \frametype[n+1][\lbrack 0,p-1 \rbrack] \times \U \\
  \end{array}
\end{equation*}

@@

The types $\paintingtype[n][\lbrack 0,p \rbrack]$ of $\painting[n][\lbrack 0,p \rbrack]$ for $p \leq n$
are type families over frames, so $\framep[n][\lbrack 0,p \rbrack]$ needs to be assumed
beforehand:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \paintingtype[n][\lbrack 0,p \rbrack] & (\framep[n][\lbrack 0,p \rbrack]:\frametype[n][\lbrack 0,p \rbrack]) & :      & \U                                                                                                                   \\
    \paintingtype[n][]                    & \unitpoint                                                           & \defeq & \unittype                                                                                                            \\
    \paintingtype[n][\lbrack 0,p \rbrack] & (\framep[n][\lbrack 0,p \rbrack])                                    & \defeq & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack]) \times (\framep[n][p] \rightarrow \U) \\
  \end{array}
\end{equation*}

@@

Assuming $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\painting[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$, we can
mutually define the types $\restrftype[n][\lbrack 0,p-1 \rbrack]$ of frame
restrictions together with $\framep[n+1][0,...p]$ as a function from
$\restrf[n][\lbrack 0,p-1 \rbrack]$, introducing also abbreviations for the types
and definitions of layers, as well as for the combination of sequences
of $\framep[n][\lbrack 0,p-1 \rbrack]$, $\painting[n][\lbrack 0,p-1 \rbrack]$ and
$\restrf[n][\lbrack 0,p-1 \rbrack]$ in one block, which we call
$\depstype[n][0,...p-1]$:

@@

From which we can deduce $\painting[n+1][\lbrack 0,p \rbrack]$ for $p \leq n+1$ by a second recursion, this time on $p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][\lbrack 0,p \rbrack]   &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \paintingtype[n+1][\lbrack 0,p \rbrack](\framep[n+1][\lbrack 0,p \rbrack](\fulldeps.\deps[n-1][\lbrack 0,p \rbrack])) \\
    \painting[n+1][\lbrack 0,0-1 \rbrack] &
    \fulldeps[n]                          & \defeq & \unittype                                                                                                             \\
    \painting[n+1][\lbrack 0,p \rbrack]   &
    \fulldeps[n]
                                          & \defeq &
    (\painting[n+1][\lbrack 0,p-1 \rbrack](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n]))                                                                                \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack]     &
    (\fulldeps[n]:\fulldepstype[n])           & :      & \U                                                                \\
    \restrptype[n][\lbrack 0,0-1 \rbrack]     & E      & \defeq                                                & \unittype \\
    \restrptype[n][\lbrack 0,(p+1)-1 \rbrack] &
    (\fulldeps[n])
                                              & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\Pi \epsilon.\Pi d:\framep[n+1][p](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                              \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                                \\
  \end{array}
\end{equation*}

@@

              \fulldeps[n][\lbrack 0,n \rbrack] & : & \fulldepstype[n][\lbrack 0,n \rbrack]               \\

@@

    \cohftype[n][\lbrack 0,0-1 \rbrack]     &

@@

    \cohftype[n][\lbrack 0,(p+1)-1 \rbrack] &

@@

From which we can deduce $\restrp[n+1][\lbrack 0,p \rbrack]$ for $p \leq n+1$ again by a second recursion, this time on $p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][\lbrack 0,p \rbrack]       &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \restrptype[n+1][\lbrack 0,p \rbrack](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1][\lbrack 0,0-1 \rbrack]     &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                    \\
    \restrp[n+1][\lbrack 0,p \rbrack]       &
    \fulldepscohs[n]
                                            & \defeq &
    (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                       \\
  \end{array}
\end{equation*}

@@

Thanks to $\cohptype[n][\lbrack 0,p-1 \rbrack]$, we can mutually define $\cohf[n+1][\lbrack 0,p \rbrack]$ and $\cohttype[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ from $\fulldepscohs[n][\lbrack 0,p-1 \rbrack]$, abbreviating this and the dependency in a $\coht[n][\lbrack 0,p-1 \rbrack]$ as $\depscohttype[n][\lbrack 0,p-1 \rbrack]$, and using also abbreviations in passing for $\cohl$ and $\cohltype$. It is also convenient to build $\depscohs[n+1][\lbrack 0,p \rbrack]$ from $\depscoht[n][\lbrack 0,p-1 \rbrack]$, so that we can now build $\framep[n+3][p]$. Thus, we also include this abbreviation in the definition:

@@

    \cohttype[n][\lbrack 0,0-1 \rbrack]                                                      &

@@

    \cohttype[n][\lbrack 0,(p+1)-1 \rbrack]                                                  &

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][p]                           &
    (\fulldepscoht[n]:\fulldepscohttype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d.\,\Pi c.\,   \\
        \restrp[n][p][q][\epsilon](\overline{\fulldeps}(\depscoht))(\restrp[n+1][p][r][\omega]
        (\fulldepscoht.\depscoht.\fulldepscohs[n+1][\lbrack 0,n \rbrack])(d)(c)) = \\
        \restrp[n][p][r][\omega](\overline{\fulldeps}(\depscoht))(\restrp[n+1][p][q+1][\epsilon]
        (\fulldepscoht.\depscoht.\fulldepscohs[n+1][\lbrack 0,n \rbrack])(d)(c))   \\
        (\fulldepscoht.\depscoht.\fulldepscohs[n+1][\lbrack 0,n \rbrack])(d)(c))   \\
      \end{array}\right)                                                                         \\
    \cohp[n+1][n]                           &
    (\fulldepscoht[n])
                                            & \defeq & \lambda q\leq n.\lambda r\leq q.\lambda \epsilon.\lambda \omega.\lambda d. \lambda (l,c). l\epsilon \\
    \cohp[n+1][p](p<n)                      &
    (\fulldepscoht[n])
                                            & \defeq & \lambda q\leq n.\lambda r\leq q.\lambda \epsilon.\lambda \omega.\lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \cohl[n][p][q][r][\epsilon][\omega](\fulldepscoht.\depscoht[n][\lbrack 0,p-1 \rbrack])(d)(l), \\
              \cohp[n+1][p+1][q][r][\epsilon][\omega](\fulldepscoht[n])(d,l))(c)                            \\
            \end{array}\right)                                                    \\
  \end{array}
\end{equation*}

From which we can deduce $\cohp[n+1][\lbrack 0,p \rbrack]$ again by a second recursion, this time on $p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][\lbrack 0,p \rbrack]         &
    (\fulldepscoht[n]:\fulldepscohttype[n]) & :      & \cohptype[n+1][\lbrack 0,p \rbrack](\fulldepscoht.\depscoht.\fulldepscohs[n])) \\
    \cohp[n+1][\lbrack 0,0-1 \rbrack]       &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                      \\
    \cohp[n+1][\lbrack 0,p \rbrack]         &
    \fulldepscoht[n]
                                            & \defeq &
    (\cohp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscoht[n]),\cohp[n+1][p](\fulldepscoht[n]))                                             \\
  \end{array}
\end{equation*}

>>>>

\newcommandx{\deps}[2][1,2]{\depsp{}[][][][#1][#2][][]}
\newcommandx{\fulldeps}[1][1]{\depsp{full}[][][][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\depsp{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohs}[1][1]{\depsp{fullcohs}[][][][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\depsp{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscoht}[1][1]{\depsp{fullcoht}[][][][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\depstypep{}[][][][#1][#2][][]}
\newcommandx{\fulldepstype}[1][1]{\depstypep{full}[][][][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\depstypep{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohstype}[1][1]{\depstypep{fullcohs}[][][][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\depstypep{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscohttype}[1][1]{\depstypep{fullcoht}[][][][#1][][][]}

@@

Assuming $\framep[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ only requires to define the type of frames, which are just finite products of $\U$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \frametype[n][\lbrack 0,p-1 \rbrack] &  & :      & \U                                               \\
    \frametype[n][\lbrack \rbrack]       &  & \defeq & \unittype                                        \\
    \frametype[n][\lbrack 0,p \rbrack]   &  & \defeq & \frametype[n+1][\lbrack 0,p-1 \rbrack] \times \U \\
  \end{array}
\end{equation*}

@@

The types $\paintingtype[n][\lbrack 0,p-1 \rbrack]$ of $\painting[n][\lbrack 0,p \rbrack]$ for $p \leq n+1$ are type families over frames, so $\framep[n][\lbrack 0,p-1 \rbrack]$ needs to be assumed beforehand:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \paintingtype[n][\lbrack 0,p-1 \rbrack] & (\framep[n][\lbrack 0,p \rbrack]:\frametype[n][\lbrack 0,p-1 \rbrack]) & :      & \U                                                                                                                   \\
    \paintingtype[n][\lbrack \rbrack]       & \unitpoint                                                             & \defeq & \unittype                                                                                                            \\
    \paintingtype[n][\lbrack 0,p \rbrack]   & (\framep[n][\lbrack 0,p \rbrack])                                      & \defeq & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack]) \times (\framep[n][p] \rightarrow \U) \\
  \end{array}
\end{equation*}

@@

Assuming $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\painting[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$, we can
mutually define the types $\restrftype[n][\lbrack 0,p-1 \rbrack]$ of frame
restrictions together with $\framep[n+1][\lbrack 0,p-1 \rbrack]$ as a function from
$\restrf[n][\lbrack 0,p-1 \rbrack]$, introducing also abbreviations for the types
and definitions of layers, as well as for the combination of sequences
of $\framep[n][\lbrack 0,p-1 \rbrack]$, $\painting[n][\lbrack 0,p-1 \rbrack]$ and
$\restrf[n][\lbrack 0,p-1 \rbrack]$ in one block, which we call
$\depstype[n][0,...p-1]$:

@@

From which we can deduce $\painting[n+1][\lbrack 0,p-1 \rbrack]$ for $p \leq n+2$ by a second recursion, this time on $p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][\lbrack 0,p-1 \rbrack] &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack])) \\
    \painting[n+1][\lbrack \rbrack]       &
    \fulldeps[n]                          & \defeq & \unittype                                                                                                                 \\
    \painting[n+1][\lbrack 0,p \rbrack]   &
    \fulldeps[n]
                                          & \defeq &
    (\painting[n+1][\lbrack 0,p-1 \rbrack](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n]))                                                                                    \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack] &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \U                                                                \\
    \restrptype[n][\lbrack \rbrack]       & E      & \defeq                                                & \unittype \\
    \restrptype[n][\lbrack 0,p \rbrack]   &
    (\fulldeps[n])
                                          & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\Pi \epsilon.\Pi d:\framep[n+1][p](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                              \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                            \\
  \end{array}
\end{equation*}

@@

              \fulldeps[n]                    & : & \fulldepstype[n]                                    \\

@@

    \cohftype[n][\lbrack \rbrack]           &

@@

    \cohftype[n][\lbrack 0,p \rbrack]       &

@@

From which we can deduce $\restrp[n+1][\lbrack 0,p-1 \rbrack]$ for $p \leq n+2$ again by a second recursion, this time on $p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][\lbrack 0,p-1 \rbrack]     &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \restrptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1][\lbrack \rbrack]           &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                      \\
    \restrp[n+1][\lbrack 0,p \rbrack]       &
    \fulldepscohs[n]
                                            & \defeq &
    (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                         \\
  \end{array}
\end{equation*}

@@

Thanks to $\cohptype[n][\lbrack 0,p-1 \rbrack]$, we can mutually define $\cohf[n+1][\lbrack 0,p \rbrack]$ and $\cohttype[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ from $\fulldepscohs[n]$, abbreviating this and the dependency in a $\coht[n][\lbrack 0,p-1 \rbrack]$ as $\depscohttype[n][\lbrack 0,p-1 \rbrack]$, and using also abbreviations in passing for $\cohl$ and $\cohltype$. It is also convenient to build $\depscohs[n+1][\lbrack 0,p \rbrack]$ from $\depscoht[n][\lbrack 0,p-1 \rbrack]$, so that we can now build $\framep[n+3][p]$. Thus, we also include this abbreviation in the definition:

@@

    \cohttype[n][\lbrack \rbrack]                                                            &

@@

    \cohttype[n][\lbrack 0,p \rbrack]                                                        &

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][p]                           &
    (\fulldepscoht[n]:\fulldepscohttype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d.\,\Pi c.\, \\
        \restrp[n][p][q][\epsilon](\overline{\fulldeps}(\depscoht))(\restrp[n+1][p][r][\omega]
        (\fulldepscoht.\depscoht.\fulldepscohs[n+1])(d)(c)) =                    \\
        \restrp[n][p][r][\omega](\overline{\fulldeps}(\depscoht))(\restrp[n+1][p][q+1][\epsilon]
        (\fulldepscoht.\depscoht.\fulldepscohs[n+1])(d)(c))                      \\
        (\fulldepscoht.\depscoht.\fulldepscohs[n+1])(d)(c))                      \\
      \end{array}\right)                                                      \\
    \cohp[n+1][n]                           &
    (\fulldepscoht[n])
                                            & \defeq & \lambda q\leq n.\lambda r\leq q.\lambda \epsilon.\lambda \omega.\lambda d. \lambda (l,c). l\epsilon \\
    \cohp[n+1][p](p<n)                      &
    (\fulldepscoht[n])
                                            & \defeq & \lambda q\leq n.\lambda r\leq q.\lambda \epsilon.\lambda \omega.\lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \cohl[n][p][q][r][\epsilon][\omega](\fulldepscoht.\depscoht[n][\lbrack 0,p-1 \rbrack])(d)(l), \\
              \cohp[n+1][p+1][q][r][\epsilon][\omega](\fulldepscoht[n])(d,l))(c)                            \\
            \end{array}\right)                                                    \\
  \end{array}
\end{equation*}

From which we can deduce $\cohp[n+1][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ again by a second recursion, this time on $p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][\lbrack 0,p-1 \rbrack]       &
    (\fulldepscoht[n]:\fulldepscohttype[n]) & :      & \cohptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscoht.\depscoht.\fulldepscohs[n])) \\
    \cohp[n+1][\lbrack \rbrack]             &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                        \\
    \cohp[n+1][\lbrack 0,p \rbrack]         &
    \fulldepscoht[n]
                                            & \defeq &
    (\cohp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscoht[n]),\cohp[n+1][p](\fulldepscoht[n]))                                               \\
  \end{array}
\end{equation*}
